The language of generalization: Probability, vagueness, and interaction

### 2018-02-18 paper/coverletter.tex

The language of generalization is most commonly studied in predications of categories (e.g., ~\emph{Birds fly}), otherwise known as \emph{generic language}.
But different semantic types can described in generalization, including events (\emph{habitual language}; e.g., ~\emph{Mary swims after work}) and causal relationships (\emph{causal language}; e.g., ~\emph{Staring at the sun makes you go blind}).
Generalizations are ubiquitous in everyday conversation (\emph{Movie theaters are cold.}), the speech of young children (\emph{Slides aren't for grown ups!}), child-directed speech (\emph{Dogs are friendly.}), political and scientific discourse (\emph{Politicians make promises.}, \emph{Psychology experiments don't replicate.}), stereotypes (\emph{Boys are good at math.}), motivation (\emph{Big girls eat broccoli!}), and many other facets of experience.

@@

We have previously received peer-review of a similar manuscript submitted to other top-ranking journals including \emph{Psychological Review}.
Those reviews questioned how much our work advances understanding of the language of generalization and highlighted the need for comparison to other models. 
The current manuscript differs from previous submissions in two major ways.
%We show how this model naturally explains the flexibility in endorsement patterns via interlocutors' prior beliefs about the property or event in the question.
The empirical and much of the philosophical literatures have focused on generalizations about categories, or generic language.
Our previous manuscript also focused on categories; in this paper, however, we investigate three case studies: generalizations about categories, events, and causes. 
Immediately we see the generality of our formal model: We are able to explain context-sensitive endorsements across these three domains without adopting domain-specific assumptions.
Furthermore, each case study naturally lends methodological advantages to provide further tests of our model.
With event generalizations, we are able to interrogate the underlying semantic scale, and find that speakers' subjective beliefs about what is likely to be the case in the future, and not the objective truth about what has been the case in the past, is what drives endorsements. 
With causal generalizations, we are able to \emph{manipulate} (rather than merely measure) prior knowledge, and show it is in fact causally related to endorsement. 
Finally, we develop three alternative models: two linear models based on empirically-elicited statistical knowledge and one rich, cognitive model which lacks the \emph{underspecification of meaning} at the core of our hypothesis. 
These alternative models fail to account for key aspects of the context-sensitivity that is observed in the data and that our full model predicts.
These alternative models and the empirical paradigms used provide strong tests of the theoretical contribution of our paper. 
%This novel approach for generalizations in language opens up a number of interesting questions about the nature of conceptual structure and its linguistic traces, which we discuss at the end of the paper.

@@

This work will be of strong interest to psychologists, as well as linguists, philosophers, and computer scientists interested in language and interaction.
The puzzles of generalizations in language are immediately compelling to a general audience, the more so because of their connections to cognitive development, cultural transmission of knowledge, and propagation of stereotypes.
Our solution represents the first formal model of generalization in language to be validated by quantitative agreement with human behavioral data.
It is not, however, uncontroversial: we champion a statistical view that has been often dismissed in the literature.
Indeed, in presentations around the world we have found this work to excite broad interest and engagement.

The following reviewers have expertise in the phenomenology of generic language, experimental studies of pragmatics, and/or the Bayesian modeling tools we use:

>>>>

The language of generalization is most commonly studied in predications of categories (e.g., ~\emph{Birds fly}), otherwise known as \emph{generic language}.
But different semantic types can be described in generalization, including events (\emph{habitual language}; e.g., ~\emph{Mary swims after work}) and causal relationships (\emph{causal language}; e.g., ~\emph{Staring at the sun makes you go blind}).
Generalizations are ubiquitous in everyday conversation (\emph{Movie theaters are cold.}), the speech of young children (\emph{Slides aren't for grown ups!}), child-directed speech (\emph{Dogs are friendly.}), political and scientific discourse (\emph{Politicians make promises.}, \emph{Psychology experiments don't replicate.}), stereotypes (\emph{Boys are good at math.}), motivation (\emph{Big girls eat broccoli!}), and many other facets of experience.

@@

We have previously received peer-review of a related manuscript submitted to top-ranking journals including \emph{Psychological Review}.
Those reviews raised a umber of points regarding the presentation and generality of our work at the time, and highlighted the need for comparison to other models. 
The current manuscript differs from previous submissions a number of important ways.
First, we have completely re-written the paper for clarity and focus, especially simplifying and streamlining the modeling sections.
Second, we have extended from generalizations about categories, or generic language, to investigate three case studies: generalizations about categories, events, and causes.
We are able to explain endorsements across these three domains without adopting domain-specific assumptions. 
Furthermore, each case study naturally lends methodological advantages to provide further tests of our model. (E.g. with causal generalizations, we are able to \emph{manipulate} (rather than merely measure) prior knowledge.)
%The empirical and much of the philosophical literatures have focused on generalizations about categories, or generic language.
%Our previous manuscript also focused on categories; in this paper, however, we investigate three case studies: generalizations about categories, events, and causes. 
%Immediately we see the generality of our formal model: We are able to explain context-sensitive endorsements across these three domains without adopting domain-specific assumptions.
%Furthermore, each case study naturally lends methodological advantages to provide further tests of our model.
%With event generalizations, we are able to interrogate the underlying semantic scale, and find that speakers' subjective beliefs about what is likely to be the case in the future, and not the objective truth about what has been the case in the past, is what drives endorsements. 
%With causal generalizations, we are able to \emph{manipulate} (rather than merely measure) prior knowledge, and show it is in fact causally related to endorsement. 
Finally, we develop three alternative models: two linear models based on empirically-elicited statistical knowledge and one rich, cognitive model which lacks the \emph{underspecification of meaning} at the core of our hypothesis. 
These alternative models fail to account for key aspects of the context-sensitivity that is observed in the data and that our full model predicts.
%These alternative models and the empirical paradigms used provide strong tests of the theoretical contribution of our paper. 
%This novel approach for generalizations in language opens up a number of interesting questions about the nature of conceptual structure and its linguistic traces, which we discuss at the end of the paper.
Overall we feel that this paper represents a much improved, generalized and streamlined, study of generalization in language.
(We would be happy to provide detailed responses to previous reviewers, however we feel this would be of limited use given the major changes in content and presentation.)

@@

This work will be of interest to psychologists, as well as linguists, philosophers, and computer scientists interested in language and interaction.
The puzzles of generalizations in language are immediately compelling to a general audience, the more so because of their connections to cognitive development, cultural transmission of knowledge, and propagation of stereotypes.
Our solution represents the first formal model of generalization in language to be validated by quantitative agreement with human behavioral data.
It is not, however, uncontroversial: we champion a statistical view that has been often dismissed in the literature.
Indeed, in presentations around the world we have found this work to excite broad interest and engagement.

We recommend the following reviewers have expertise in the phenomenology of generic language, experimental studies of language use, and/or the Bayesian modeling tools we use:
